\chapter{Real Analysis}

Real analysis provides the rigorous foundation for calculus and studies the properties of real numbers, sequences, series, and functions.

\section{The Real Number System}

\begin{definition}[Supremum]
Let $S \subseteq \mathbb{R}$ be bounded above. The \textbf{supremum} (or least upper bound) of $S$, denoted $\sup S$, is the smallest number that is greater than or equal to every element of $S$.
\end{definition}

\begin{theorem}[Completeness Axiom]
Every non-empty subset of $\mathbb{R}$ that is bounded above has a supremum in $\mathbb{R}$.
\end{theorem}

\begin{definition}[Archimedean Property]
For any real numbers $x$ and $y$ with $x > 0$, there exists a positive integer $n$ such that $nx > y$.
\end{definition}

\section{Sequences and Series}

\begin{definition}[Convergent Sequence]
A sequence $(a_n)$ \textbf{converges} to $L$ if for every $\varepsilon > 0$, there exists $N \in \mathbb{N}$ such that for all $n \geq N$:
\[|a_n - L| < \varepsilon\]
We write $\lim_{n \to \infty} a_n = L$.
\end{definition}

\begin{theorem}[Monotone Convergence Theorem]
Every bounded monotone sequence converges.
\end{theorem}

\begin{proof}
Let $(a_n)$ be a bounded increasing sequence. Since $(a_n)$ is bounded above, by the completeness axiom, $S = \{a_n : n \in \mathbb{N}\}$ has a supremum $L = \sup S$.

For any $\varepsilon > 0$, since $L$ is the least upper bound, $L - \varepsilon$ is not an upper bound. Therefore, there exists $N$ such that $a_N > L - \varepsilon$.

Since $(a_n)$ is increasing, for all $n \geq N$:
\[L - \varepsilon < a_N \leq a_n \leq L < L + \varepsilon\]

Thus $|a_n - L| < \varepsilon$, proving convergence to $L$.
\end{proof}

\begin{definition}[Cauchy Sequence]
A sequence $(a_n)$ is \textbf{Cauchy} if for every $\varepsilon > 0$, there exists $N \in \mathbb{N}$ such that for all $m, n \geq N$:
\[|a_m - a_n| < \varepsilon\]
\end{definition}

\begin{theorem}[Cauchy Criterion]
A sequence in $\mathbb{R}$ converges if and only if it is Cauchy.
\end{theorem}

\section{Continuity and Uniform Continuity}

\begin{definition}[Continuity at a Point]
A function $f: D \to \mathbb{R}$ is \textbf{continuous at} $c \in D$ if for every $\varepsilon > 0$, there exists $\delta > 0$ such that for all $x \in D$:
\[|x - c| < \delta \implies |f(x) - f(c)| < \varepsilon\]
\end{definition}

\begin{definition}[Uniform Continuity]
A function $f: D \to \mathbb{R}$ is \textbf{uniformly continuous} on $D$ if for every $\varepsilon > 0$, there exists $\delta > 0$ such that for all $x, y \in D$:
\[|x - y| < \delta \implies |f(x) - f(y)| < \varepsilon\]
\end{definition}

\begin{theorem}[Uniform Continuity on Compact Sets]
If $f$ is continuous on a compact set $K$, then $f$ is uniformly continuous on $K$.
\end{theorem}

\section{Differentiation}

\begin{theorem}[Rolle's Theorem]
If $f$ is continuous on $[a,b]$, differentiable on $(a,b)$, and $f(a) = f(b)$, then there exists $c \in (a,b)$ such that $f'(c) = 0$.
\end{theorem}

\begin{theorem}[Mean Value Theorem]
If $f$ is continuous on $[a,b]$ and differentiable on $(a,b)$, then there exists $c \in (a,b)$ such that:
\[f'(c) = \frac{f(b) - f(a)}{b - a}\]
\end{theorem}

\begin{definition}[Uniform Convergence]
A sequence of functions $(f_n)$ \textbf{converges uniformly} to $f$ on a set $E$ if for every $\varepsilon > 0$, there exists $N \in \mathbb{N}$ such that for all $n \geq N$ and all $x \in E$:
\[|f_n(x) - f(x)| < \varepsilon\]
\end{definition}

\begin{theorem}[Uniform Convergence and Continuity]
If $(f_n)$ is a sequence of continuous functions that converges uniformly to $f$ on a set $E$, then $f$ is continuous on $E$.
\end{theorem}

\section{Integration}

\begin{definition}[Riemann Integrable]
A bounded function $f$ on $[a,b]$ is \textbf{Riemann integrable} if:
\[\overline{\int_a^b} f = \underline{\int_a^b} f\]
where $\overline{\int_a^b} f$ and $\underline{\int_a^b} f$ are the upper and lower Darboux integrals, respectively.
\end{definition}

\begin{theorem}[Riemann Integrability Criterion]
A bounded function $f$ on $[a,b]$ is Riemann integrable if and only if the set of discontinuities of $f$ has measure zero.
\end{theorem}

\begin{theorem}[Fundamental Theorem of Calculus (Rigorous)]
Let $f$ be Riemann integrable on $[a,b]$.
\begin{enumerate}
    \item If $F(x) = \int_a^x f(t) \, dt$, then $F$ is continuous on $[a,b]$.
    \item If $f$ is continuous at $c \in [a,b]$, then $F'(c) = f(c)$.
    \item If $g$ is differentiable on $[a,b]$ and $g' = f$, then $\int_a^b f(x) \, dx = g(b) - g(a)$.
\end{enumerate}
\end{theorem}

\section{Metric Spaces}

\begin{definition}[Metric Space]
A \textbf{metric space} is a set $X$ together with a function $d: X \times X \to [0, \infty)$ such that for all $x, y, z \in X$:
\begin{enumerate}
    \item $d(x, y) = 0$ if and only if $x = y$
    \item $d(x, y) = d(y, x)$ (symmetry)
    \item $d(x, z) \leq d(x, y) + d(y, z)$ (triangle inequality)
\end{enumerate}
\end{definition}

\begin{definition}[Open Ball]
In a metric space $(X, d)$, the \textbf{open ball} of radius $r > 0$ centered at $x \in X$ is:
\[B_r(x) = \{y \in X : d(x, y) < r\}\]
\end{definition}

\begin{definition}[Compact Set]
A subset $K$ of a metric space is \textbf{compact} if every open cover of $K$ has a finite subcover.
\end{definition}

\begin{theorem}[Heine-Borel Theorem]
A subset of $\mathbb{R}^n$ is compact if and only if it is closed and bounded.
\end{theorem}

\section{Exercises}

\begin{enumerate}
    \item Prove that $\lim_{n \to \infty} \frac{1}{n} = 0$ using the definition of convergence.
    
    \item Show that the sequence $a_n = \frac{n}{n+1}$ is Cauchy.
    
    \item Prove that $f(x) = x^2$ is uniformly continuous on any bounded interval $[a,b]$.
    
    \item Use the Mean Value Theorem to prove that $|\sin x - \sin y| \leq |x - y|$ for all real $x, y$.
    
    \item Show that the function $f(x) = \frac{1}{x}$ is not uniformly continuous on $(0,1)$.
\end{enumerate}
